Dendritic Ca2+ as a Predictor of Stimulus Perception and Behavior

### 2017-08-01 presentation/demo.tex
asd

\title{Uni- and Multivariate Analysis of Dendritic Ca2+ Data}
\subtitle{In a Stimulus Detection Task}
\date{\today}
%\date{}
\author{Georg Chechelnizki}
\institute{BCCN Berlin}
% \titlegraphic{\hfill\includegraphics[height=1.5cm]{logo.pdf}}

@@

\begin{frame}{Table of contents}
  \setbeamertemplate{section in toc}[sections numbered]
  \tableofcontents[hideallsubsections]
\end{frame}

@@

\section{Goal}
\begin{frame}[fragile]{Goal}
The goal of this project is to investigate the following:
\begin{itemize}
\only<+->{
\item What is the relationship between Ca2+ activity and stimulus intensity/behavior}
\only<+->{
\item Can we build good classifiers for this data?}
\only<+->{
\item Does a multivariate approach give us any significant advantage over a univariate one?}
\end{itemize}
\end{frame}

@@

@@

\begin{frame}[fragile]{$\text{Ca}^{2+}$-Trace of Largest $|\text{DI}|$-Dendrites}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{on_vs_off.png}
      \caption*{Near threshold-stimulus ($\approx 1$)}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{SVM - What is an SMV?}
\begin{center}
\only<+->{
\begin{figure}
      \includegraphics[height=0.3\textwidth]{svm_exp1.png}
	\end{figure}}
	%\only<+->{
	%\begin{figure}
    %  \includegraphics[height=0.3\textwidth]{svm_exp.png}
	%\end{figure}}
	\end{center}
	\begin{columns}[T,onlytextwidth]
    \column{0.5\textwidth}
    	\only<+->{
      \begin{figure}
      \includegraphics[height=0.7\textwidth]{svm_exp2.png}
      \end{figure}}

    \column{0.5\textwidth}
		\only<+->{
      \begin{figure}
      \includegraphics[height=0.7\textwidth]{svm_exp3.png}
      \end{figure}}

\begin{frame}[fragile]{Why SVM?}

\begin{itemize}
\item staple of machine learning
\item regularization parameter limits overfitting
\item good for small samples with many features (in theory)
\end{itemize}

\begin{frame}[fragile]{SVM Specifics}
\begin{itemize}
\item All data are normalized to zero mean and unit variance
\item Crossvalidation is performed to conrol for overfitting
\end{itemize}

\begin{frame}[fragile]{SVM - Most Accurate Dendrites}
\begin{table}
    \caption*{Stimulus strength 1 (near thereshold)}
    \begin{tabular}{c|c|c}
      \toprule
      Dendrite \# & $\mu_{acc}$ & $\sigma_{acc}$\\
      \midrule
      57 & 0.70 & 0.12\\
      53 & 0.68 & 0.14\\
      48 & 0.63 & 0.13\\
      27 & 0.62 & 0.08\\
      \bottomrule
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}[fragile]{SVM - Numbers of Classifying Dendrites by Stimulus}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{n_acc.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}{Weighted Rank Order Correlation Coefficient}
One possible soulution:
$$
\rho_W(R,Q) = 1-\frac{-2\sum_i^nw_iD_i^2}{\sum_i^nw_i(n-2i+1)^2}
$$
With a monotonicity constraint on $W$.
\end{frame}

\begin{frame}[fragile]{Rank Order Correlations of Dendrites Over Stimuli}
\begin{center}
	\begin{figure}\caption*{Weighted rank coefficient $\rho_{\omega}$ (todo: find more optimal weights)}
      \includegraphics[width=1.0\textwidth]{rank_presence.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Tuning Curves}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{tuning_ca2+single.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Tuning Curves}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{tuning_ca2+.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Bevahior Detection}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{data_hm.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}[fragile]{Behavorial Predition - Imbalance}
\begin{center}
	\begin{figure}
	  \caption*{$N_{no-lick}$ vs. $N_{lick}$ for all stimuli}
      \includegraphics[width=1.0\textwidth]{fraction.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}[fragile]{Rank Correlations of Behavoreal and Presence Dendrites}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{transrank.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Timing - What if we use different Averaging Windows?}
So far we have always averaged over the first second after the stimulus. \newline

What if we average over the second before or the second second after the stimulus?
\end{frame}

\begin{frame}[fragile]{Statistical Significance - Stimulus Presence}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{statsig.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}[fragile]{Presence Detection - Feature Selection}
\only<+->{
Since we have only few datapoints per condition (30-50 per class) in comparison to the number of features (100-150), we are prone to overfit to noise.}

@@

\only<+->{
We can use the previously best discriminating dendrites (SVM/DI).}
\end{frame}

\begin{frame}[fragile]{SVM Performance on Combined Dendrites - Presence Detection}
\begin{center}
	\begin{figure}
	%\caption*{todo: tens, improve weights}
      \includegraphics[width=1.0\textwidth]{combined_presence.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}[fragile]{Population Coding - Idea (Mante, Sussillo 2013)}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{trafo.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}[fragile]{Population Response in Task Variable Space}
\begin{center}
	\begin{figure}
	\caption*{Solve Problem with combined SVM first}
      \includegraphics[width=0.7\textwidth]{reg.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Population Response in Task Variable Space}
\begin{center}
	\begin{figure}
	\caption*{Solve Problem with combined SVM first}
      \includegraphics[width=0.7\textwidth]{reg_sum.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\section*{Extra Slides}
\begin{frame}[fragile]{Weighted Rank Order Correlation Coefficient}
\only<+->{
We want to quantify the similarity between two rank orders of dendrites $R$ and $Q$, which both have length $n$.}

>>>>

\title{Dendritic Ca2+ as a Predictor of Stimulus Perception and Behavior}
\subtitle{Uni- and Multivariate Analysis}
\date{\today}
%\date{}
\author{Georg Chechelnizki}
\institute{BCCN Berlin}
% \titlegraphic{\hfill\includegraphics[height=1.5cm]{logo.pdf}}

@@

%\begin{frame}{Table of contents}
%  \setbeamertemplate{section in toc}[sections numbered]
%  \tableofcontents[hideallsubsections]
%\end{frame}

@@

\section{Goal}
\begin{frame}[fragile]{Goal}
The goal of this project is to investigate the following:
\begin{itemize}
\only<+->{
\item What is the relationship between Ca2+ activity and stimulus intensity/behavior}
\only<+->{
\item Can we build predictive classifiers for this data?}
\only<+->{
\item Does a multivariate approach give us any significant advantage over a univariate one?}
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{SVM - What is an SVM?}
\begin{columns}[T,onlytextwidth]
\column{0.5\textwidth}
\only<1->{
\begin{center}
\begin{figure}
      \includegraphics[height=0.6\textwidth]{svm_exp1.png}
	\end{figure}}
	%\only<+->{
	%\begin{figure}
    %  \includegraphics[height=0.3\textwidth]{svm_exp.png}
	%\end{figure}}
	\end{center}
\column{0.5\textwidth}
\only<4->{
SMVs...
\begin{itemize}
\item are a staple of machine learning
\item have a regularization parameter that can combat overfitting
\item are good for small samples with many features (in theory)
\end{itemize}
}
\end{columns}
	\begin{columns}[T,onlytextwidth]
    \column{0.5\textwidth}
    	\only<2->{
      \begin{figure}
      \includegraphics[height=0.7\textwidth]{svm_exp2.png}
      \end{figure}}

    \column{0.5\textwidth}
		\only<3->{
      \begin{figure}
      \includegraphics[height=0.7\textwidth]{svm_exp3.png}
      \end{figure}}

\begin{frame}[fragile]{SVM Specifics}
\begin{itemize}
\only<+->{
\item All data are normalized to zero mean and unit variance}
\only<+->{
\item Crossvalidation is performed to conrol for overfitting}
\end{itemize}

@@

\begin{frame}[fragile]{SVM - Most Accurate Dendrites}
%\begin{table}
%    \caption*{Stimulus strength 1 (near thereshold)}
%    \begin{tabular}{c|c|c}
%      \toprule
%      Dendrite \# & $\mu_{acc}$ & $\sigma_{acc}$\\
%      \midrule
%      57 & 0.70 & 0.12\\
%      53 & 0.68 & 0.14\\
%      48 & 0.63 & 0.13\\
%      27 & 0.62 & 0.08\\
%      \bottomrule
%    \end{tabular}
%  \end{table}

\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{presence_single.png}
	\end{figure}
	\end{center}
\end{frame}

%\begin{frame}[fragile]{SVM - Numbers of Classifying Dendrites by Stimulus}
%\begin{center}%
%	\begin{figure}
%      \includegraphics[width=1.0\textwidth]{n_acc.png}
%	\end{figure}
%	\end{center}
%\end{frame}

\begin{frame}[fragile]{SVM - Statistical Significance}
\begin{center}
\only<+->{
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{statsig_0.png}
	\end{figure}}
\only<+->{
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{statsig_1.png}
	\end{figure}}
	\end{center}
\end{frame}

@@

\only<+->{
\begin{table}
    \begin{tabular}{c|c|c|c|c}
      \toprule
      Dendrite & $R_i$ & $Q_i$ & $R_{acc}$ & $Q_{acc}$\\
      \midrule
      1 & 1 & 2 & 0.78 & 0.81\\
      2 & 2 & 1 & 0.72 & 0.76\\
      3 & 3 & 3 & 0.65 & 0.70\\
      4 & 4 & 4 & 0.58 & 0.57\\
      5 & 5 & 16 & 0.51 & 0.53\\
      6 & 6 & 75 & 0.53 & 0.48\\
      7 & 7 & 8 & 0.51 & 0.52\\
      8 & 8 & 89 & 0.52 & 0.49\\
      9 & 9 & 100 & 0.50 & 0.51\\
      ... & ... & ... & ... & ...

    \end{tabular}
  \end{table}
}
\end{frame}

\begin{frame}[fragile]{Rank Order Correlation}
\only<+->{
Standard approach: \textbf{Spearman's rank order coefficient} 
$\boldsymbol{\rho}$}

\only<+->{
$$
\rho(R,Q) = 1 - \frac{6\sum_i^nD_i^2}{n(n	^2-1)}
$$
}

\only<+->{
Problem: Many of the dendrites are not significant for classification but affect $\rho$ greatly. We would like to be able to \textbf{weight our} $\mathbf{R}$'s and $\mathbf{Q}$'s.}
\end{frame}

\begin{frame}{Weighted Rank Order Correlation Coefficient}
One possible solution:
$$
\rho_W(R,Q) = 1-\frac{-2\sum_i^nw_iD_i^2}{\sum_i^nw_i(n-2i+1)^2}
$$
With a monotonicity constraint on $W$.
\end{frame}

\begin{frame}[fragile]{Rank Order Correlations of Dendrites Over Stimuli}

	\begin{columns}[T,onlytextwidth]
    \column{0.5\textwidth}
    \begin{center}
    \only<+->{
	\begin{figure}
	\caption*{Experiment 1}
      \includegraphics[width=1.0\textwidth]{rank_presence_0.png}
	\end{figure}}
	\end{center}
	\column{0.5\textwidth}
	\begin{center}
	\only<+->{
	\begin{figure}
	\caption*{Experiment 2}
      \includegraphics[width=1.0\textwidth]{rank_presence.png}
	\end{figure}}
	\end{center}
	\end{columns}
\end{frame}

\begin{frame}[fragile]{Responses of One Dendrite to Different Stimuli}

	\begin{columns}[T,onlytextwidth]
    \column{0.5\textwidth}
    \begin{center}
    \only<+->{
	\begin{figure}
	\caption*{Experiment 1}
      \includegraphics[width=1.0\textwidth]{tuning_ca2+_single_0.png}
	\end{figure}}
	\end{center}
	\column{0.5\textwidth}
	\begin{center}
	\only<+->{
	\begin{figure}
	\caption*{Experiment 2}
      \includegraphics[width=1.3\textwidth]{tuning_ca2+single.png}
	\end{figure}}
	\end{center}
	\end{columns}
\end{frame}

%\begin{frame}[fragile]{Tuning Curves}
%\begin{center}
%	\begin{figure}
%      \includegraphics[width=1.0\textwidth]{tuning_ca2+single.png}
%	\end{figure}
%	\end{center}
%\end{frame}

%\begin{frame}[fragile]{Tuning Curves}
%\begin{center}
%	\begin{figure}
%      \includegraphics[width=1.0\textwidth]{tuning_ca2+.png}
%	\end{figure}
%	\end{center}
%\end{frame}

\begin{frame}[fragile]{Behavior Detection}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{data_hm.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}[fragile]{Behavioral Predition - Imbalance}
\begin{center}
	\begin{figure}
	  \caption*{$N_{no-lick}$ vs. $N_{lick}$ for all stimuli}
      \includegraphics[width=1.0\textwidth]{fraction.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}[fragile]{Rank Order Correlations of Best Behavior and Presence Dendrites}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{tranksrank_3x3.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Timing - What if We Use Different Averaging Windows?}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{timing_exp.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Timing - Statistical Significance}
\begin{center}
\only<+->{
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{statsig_t0.png}
	\end{figure}}
\only<+->{
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{statsig_t1.png}
	\end{figure}}
	\end{center}
\end{frame}

%\begin{frame}[fragile]{Statistical Significance - Stimulus Presence}
%\begin{center}
%	\begin{figure}
%      \includegraphics[width=1.0\textwidth]{statsig.png}
%	\end{figure}
%	\end{center}
%\end{frame}

@@

\begin{frame}[fragile]{Presence Detection - Feature Selection}
\only<+->{
Since we have fewer datapoints per condition (30-50 per class) than features (100-150), we are prone to overfit to noise (which is abundant in our data).}

@@

\only<+->{
We can use the previously best discriminating dendrites.}
\end{frame}

\begin{frame}[fragile]{SVM Performance on Combined Dendrites - Presence Detection, Experiment 1}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{combined_presence_0.png}
	\end{figure}
	\end{center}
\end{frame}

@@

%\begin{frame}[fragile]{Population Coding - Idea (Mante, Sussillo 2013)}
%\begin{center}
%	\begin{figure}
%      \includegraphics[width=1.0\textwidth]{trafo.png}
%	\end{figure}
%	\end{center}
%\end{frame}

@@

\begin{frame}[fragile]{Population Response in Task Variable Space}
\begin{center}
	\begin{figure}
	\caption*{Solve Problem with combined SVM first}
      \includegraphics[width=1.0\textwidth]{reg.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Population Response in Task Variable Space}
\begin{center}
	\begin{figure}
	\caption*{Solve Problem with combined SVM first}
      \includegraphics[width=1.0\textwidth]{reg_sum.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Conclusions}
\begin{itemize}
\item We can build high accuracy classifiers based on this data
\item Multivariate analysis is advantageous
\item Prestimulus Ca2+ activity does not predict the animal's decision
\item Different dendrites code for stimulus and behavior
\item There are very large differences between different animals/experiments
\item We need more data - particularily for multivariate analysis
\end{itemize}
\end{frame}

@@

\section*{Extra Slides}

\begin{frame}[fragile]{$\text{Ca}^{2+}$-Traces of Largest $|\text{AUC}-0.5|$-Dendrites}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{on_vs_off.png}
      \caption*{Near threshold-stimulus ($\approx 1$)}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{SVM Performance on Combined Dendrites - Presence Detection, Experiment 2}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{combined_presence.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Weighted Rank Order Correlation Coefficient}
\only<+->{
We want to quantify the similarity between two rank orders of dendrites $R$ and $Q$, which both have length $n$.}
